T-reX: A python package to quantify supply chain flows of waste and material in LCA databases

### 2024-03-05 latex_files/cover_letter.tex
update abstract and introduction

We are pleased to submit our original research article titled ``WasteAndMaterialFootprint: A Python Package to Quantify Supply Chain Flows of Waste and Material in LCA Databases'' for consideration by \textit{Resources, Conservation and Recycling}. This manuscript details the development and application of an innovative tool designed to improve the quantification of waste and material flows within Life Cycle Assessment (LCA) databases, an essential aspect of progressing towards sustainable resource management and circular economy objectives.

>>>>

We are pleased to submit our original research article titled ``T-reX: A Python Package to Quantify Supply Chain Flows of Waste and Material in LCA Databases'' for consideration by \textit{Resources, Conservation and Recycling}. This manuscript details the development and application of an innovative tool designed to improve the quantification of waste and material flows within Life Cycle Assessment (LCA) databases, an essential aspect of progressing towards sustainable resource management and circular economy objectives.

### 2024-03-05 latex_files/introduction.tex
update abstract and introduction

Thus, we were motivated to create a flexible and policy-driven approach to understanding a product`s demand on the waste management system both now and in the future to ensure that models of circularity and improvements in technology and policy for waste are included in LCA modeling.

\subsection{LCA}

@@

Several LCIA methods include, to some extent, waste generation \citep{foen2021ecofactors,hauschild2003edip,cen2019en15804} and material consumption \citep{arvidsson2020csi,foen2021ecofactors}. These methods, however, are generally limited in their scope (especially for waste), do not allow for flexible quantification of specific waste and material types, and often provide results in characterised units that are abstract or difficult to interpret (e.g., kg-Si equivalents or Ümweltbelastungspunkte (UBP)).

\subsection{Material Demand in LCA}
In the context of a mineral-hungry renewable energy transition and recent geo-political tensions, more attention is being paid to the security of supply of materials, especially those considered `critical raw materials' (CRMs) \citep{eu2023crmstudy,hool2023crm,mancini2013supplysecurity,jrc2023supplychain,hartley2024cepolitics,salviulo2021supplychain}. Given the increasing focus placed on this, as well as ideals of improving resource efficiency and developing the circular economy, it is essential to understand the material demands of products and processes. While LCA seeks to model the technosphere (the sum of all supply chains), its focus is often on the environmental impacts of the system---the endpoints--- rather than the primary material flows themselves.

A relatively new method, termed the crustal scarcity indicator (CSI)~\citep{arvidsson2020csi}, was developed to assess long-term global scarcity of minerals in life cycle assessment (LCA). This method introduced crustal scarcity potentials (CSPs) measured in kg silicon equivalents per kg element, derived from crustal concentrations. CSPs, provided for 76 elements, reflect the long-term global elemental scarcity based on crustal concentration proxies. The CSI, calculated by multiplying CSPs with extracted masses, effectively gauges the impact of elemental extraction.

While useful for its stated purpose, the CSI presents its midpoint results in an abstract unit (kg-Si eq.) that is difficult to interpret and compare with other impact categories. Furthermore, the CSPs are not available for all elements, and the method does not allow for the quantification of material demands in terms of mass or volume.

@@

Though often described simply as `a material with a negative economic value' \citep{guinee2004economicallocation}, waste is a nebulous concept, and one whose definition is poorly delineated and variable across space and time.  Moreover, from a systems perspective, the notion of waste is anathema to the circular economy, and it is far more useful to consider the identity and nature of the specific material flows. Thus, precise and detailed categorisation of these is essential to understand the `circularity' of an activity and its life cycle externalities. There is a conspicuous gap in the understanding of the waste footprint of human activities and their relationship with environmental damage \citep{laurenti2023wastefootprint}. Conventional LCAs consider waste as a `service' \citep{guinee2021wasteisnotaservice} and typically use generic waste processing models \citep{beylot2018} that break the causal link between the functional unit and the waste-associated impacts.

@@

A significant portion of a product's total waste is generated during earlier stages such as resource extraction, transportation, and manufacturing, often remaining `invisible' in traditional life cycle assessment (LCA) practices~\citep{laurenti2016wastefootprint}. This oversight in measuring and communicating a cradle-to-grave product waste footprint (PWF) highlights a gap in circular economy indicators. Traditional LCA does not typically view waste as having environmental significance by itself, focusing instead on emissions and resource use resulting from waste treatment. The environmental significance of waste and its correlation with other indicators has been the subject of extensive research. For example, studies have shown that popular resource footprints can cover a significant portion of environmental impact variance in product rankings~\citep{steinmann2017resourcefootprints}. However, correlations between various environmental indicators are not always consistent, as seen with the carbon footprint, which often does not correlate with other impact assessment scores~\citep{laurent2012carbonfootprint}. The aggregation of waste in PWFs raises concerns among LCA experts, regarding the uncertainties introduced by aggregated measures, as well as the potential misrepresentation of environmental performance due to differences in waste types~\citep{chen2021methoduncertainty,huijbregts2010energyfootprint}.

Moreover, existing LCA methodologies offer limited direct indicators at the impact assessment level, providing sparse information on the impacts of waste. This limitation becomes particularly evident when attempting to identify waste generation hotspots within a product's life cycle. Addressing these hotspots is crucial for advancing towards circularity. Thus, there is a pressing need for more comprehensive methods that can effectively quantify waste impacts and contribute to a better understanding of a product's total environmental footprint.
There is currently a lack of a convenient and flexible method for calculating waste flows in Life Cycle Assessment (LCA).

\cite{laurenti2023wastefootprint} developed a method to calculate the waste footprint of a product or service based on solving the demand vectors of the activities, also presenting simple measures to quantify waste hazardousness and circularity. In that study, it was shown that the waste footprint correlates well with other LCIA methods, particularly human health. The method presented, however, is limited in its scope and flexibility, is computationally intensive, difficult to use, is not easily reproducible, and suffers from errors due to double counting. The WMF tool presented in the current work provides a more flexible, transparent, and user-friendly approach to quantifying waste flows in LCA. Moreover, the WMF tool is not limited to waste but can be used to quantify any supply-chain flow, such as water, gas, and critical raw materials.

\subsection{The WasteAndMaterialFootprint Tool}

To better assess waste and material flows in LCA, the authors have developed a Python program built on the Brightway2 framework and designed to track these exchanges by translating them into indicators and impact categories. In this study, we present the WasteAndMaterialFootprint (WMF) tool that enables LCA practitioners to easily aggregate the mass and volume of any desired exchange, and to creation of flexible categories to differentiate between waste types and End-of-Life (EOL) handling using (in this case) the Ecoinvent 3.9 cutoff database.

The purpose of the WMF tool is not to quantify the environmental impacts of waste treatment, but rather to quantify the waste flows themselves, even those that are finally consumed by waste treatment processes. It provides, thus, not an impact assessment in the traditional sense, but an accounting of the waste generated by a product or service, regardless of the end-of-life fate of these flows. By definition, the development of the `circular economy' necessitates the reduction and ultimate elimination of waste---though whether this objective is thermodynamically impossible has long been the subject of lively debate by~\cite{ayres1998recycling},~\cite{reuter2012recyclinglimits} and many others. In any case, waste avoidance is of critical importance and by quantifying and classifying the waste total attributed to a product or service in an LCA database, the WMF tool provides a practical means to identify hotspots and opportunities for waste reduction and efficiency.

While methods with similar aims exist, they either lack flexibility and specificity \citep{foen2021ecofactors} or can be cumbersome to apply and suffer from errors due to multiple counting \citep{laurenti2023wastefootprint}.

This tool provides a method for the calculation of waste footprint impact category results, differentiated by the type of waste handling. Furthermore, the tool facilitates rapid investigation and identification of waste hotspots, enabled by standard contribution analysis and Sankey diagram visualization tools. The authors consider this a crucial step in addressing the deficit of Life Cycle Assessment (LCA) methods that consider waste flows in the evaluation of a product or process' circular economy potential.

>>>>

@@

Several LCIA methods include, to some extent, waste generation \citep{foen2021ecofactors,hauschild2003edip,cen2019en15804} and material consumption \citep{arvidsson2020csi,foen2021ecofactors}. These methods, however, are generally limited in their scope (especially for waste), do not allow for flexible quantification of specific waste and material types, and often provide results in characterised units that are abstract or difficult to interpret (e.g., Ümweltbelastungspunkte (UBP)).

In the context of a mineral-hungry renewable energy transition and recent geo-political tensions, more attention is being paid to the security of supply of materials, especially those considered `critical raw materials' (CRMs) \citep{eu2023crmstudy,hool2023crm,mancini2013supplysecurity,jrc2023supplychain,hartley2024cepolitics,salviulo2021supplychain}. While LCA seeks to model the technosphere (a.k.a. the anthroposphere), its focus is often on the environmental impacts of the system---the endpoints--- rather than the primary material flows themselves.

A relatively new method, termed the crustal scarcity indicator (CSI)~\citep{arvidsson2020csi}, was developed to assess long-term global scarcity of minerals in LCA. This method introduced crustal scarcity potentials (CSPs) measured in kg silicon equivalents per kg element, derived from crustal concentrations. CSPs, provided for 76 elements, reflect the long-term global elemental scarcity based on crustal concentration proxies. The CSI, calculated by multiplying CSPs with extracted masses, effectively gauges the impact of elemental extraction.

While useful for its stated purpose, the CSI presents its midpoint results in an abstract unit (kg-Si eq.) that is difficult to interpret and compare with other impact categories. Furthermore, the CSPs are not available for all elements (or more complex materials), and the method does not allow for the quantification of material demands in terms of mass or volume.

@@

Though often described simply as a `material with a negative economic value' \citep{guinee2004economicallocation}, waste is a nebulous concept, and one whose definition is poorly delineated and variable across space and time.  Moreover, from a systems perspective, the notion of waste is anathema to the circular economy, and it is far more useful to consider the identity and nature of the specific material flows. Thus, precise and detailed categorisation of these is essential to understand the `circularity' of an activity and its life cycle externalities. There is a conspicuous gap in the understanding of the waste footprint of human activities and their relationship with environmental damage \citep{laurenti2023wastefootprint}. Conventional LCAs consider waste as a `service' \citep{guinee2021wasteisnotaservice} and typically use generic waste processing models \citep{beylot2018} that break the causal link between the functional unit and the waste-associated impacts.

@@

A significant portion of a product's total waste is generated during earlier stages such as resource extraction, transportation, and manufacturing, often remaining `invisible' in traditional LCA practices~\citep{laurenti2016wastefootprint}. This oversight in measuring and communicating a cradle-to-grave product waste footprint (PWF) highlights a gap in circular economy indicators. Traditional LCA does not typically view waste as having environmental significance by itself, focusing instead on emissions and resource use resulting from waste treatment. The environmental significance of waste and its correlation with other indicators has been the subject of extensive research. For example, studies have shown that popular resource footprints can cover a significant portion of environmental impact variance in product rankings~\citep{steinmann2017resourcefootprints}. However, correlations between various environmental indicators are not always consistent, as seen with the carbon footprint, which often does not correlate with other impact assessment scores~\citep{laurent2012carbonfootprint}. The aggregation of waste in PWFs raises concerns among LCA experts, regarding the uncertainties introduced by aggregated measures, as well as the potential misrepresentation of environmental performance due to differences in waste types~\citep{chen2021methoduncertainty,huijbregts2010energyfootprint}.

Moreover, existing LCA methodologies offer limited direct indicators at the impact assessment level, providing sparse information on the impacts of waste. This limitation becomes particularly evident when attempting to identify waste generation hotspots within a product's life cycle. Addressing these hotspots is crucial for advancing towards circularity, however,  there is a lack of a convenient and flexible way to calculate waste flows in LCA and a pressing need for more comprehensive methods that can effectively quantify waste flows and, therefore, contribute to a better understanding of a product's total environmental footprint.
\cite{laurenti2023wastefootprint} developed a method to calculate the waste footprint of a product or service based on solving the demand vectors of the activities, also presenting simple measures to quantify waste hazardousness and circularity. In that study, it was shown that the waste footprint correlates well with other LCIA methods, particularly human health. The method presented, however, is limited in its scope and flexibility, is computationally intensive, difficult to use, is not easily reproducible, and suffers from errors due to double counting. The T-reX tool presented herein provides a more flexible, transparent, and user-friendly approach to quantifying waste flows in LCA. Moreover, the T-reX tool is not limited to waste but can be used to quantify any supply-chain flow, such as water, gas, and critical raw materials.

\subsection{The T-reX Tool}

To better assess waste and material flows in LCA, we have developed a Python program built on the Brightway framework~\citep{mutel2017brightway} and designed to track these exchanges by translating them into indicators and `pseudo' LCA impact (LCIA) categories. In this study, we present the T-reX tool that enables LCA practitioners to manipulate their databases to allow them to easily aggregate the mass and volume of any desired exchange, and to create flexible categories that differentiate between material categories, waste types, and EOL handling.

While methods with similar aims exist, they lack customisability and specificity \citep{foen2021ecofactors} or can be cumbersome to apply and suffer from errors due to multiple counting \citep{laurenti2023wastefootprint}.

The purpose of the T-reX tool is not to quantify the environmental impacts of material consumption and waste production, but rather to quantify the material and waste flows themselves, even those that are finally consumed by waste treatment processes. It provides, thus, not an impact assessment in the traditional sense, but an accounting of the material consumed and waste generated by a product or service inside of the technosphere, regardless of the end-of-life fate of these flows. By definition, the development of the `circular economy' necessitates the reduction and ultimate elimination of waste---though whether this objective is thermodynamically impossible has long been the subject of lively debate by~\cite{ayres1998recycling},~\cite{reuter2012recyclinglimits} and many others. In any case, avoiding material consumption and generation of waste is of critical importance. By allowing LCA practitioners to easily classify and quantify these exchanges, the T-reX tool provides a practical means to identify hotspots and opportunities for waste reduction and material efficiency.

### 2024-03-05 latex_files/main.tex
update abstract and introduction

    % SPOTLIGHTS (5 bullet points, max 120 characters each)
    {\Large Spotlights}
    \vspace{1em}
    \begin{description}[style=nextline]
        \item[Bullet 1: Critical context and background information on the problem addressed] Understanding waste and material flows is essential for developing the circular economy and supply chain resilience
        \item[Bullet 2: A brief overview of the key finding of the study (or findings if necessary)] A program was developed to quantify waste and material footprints of activities in life cycle assessment (LCA) databases
        \item[Bullet 3: The most radical, creative, disruptive or innovative aspect of the manuscript] Built to complement the Brightway LCA framework and ActivityBrowser, the tool is flexible and easy to use.
        \item[Bullet 4: The significance of the results to the environment, economics or society] The tool can be used to identify pre-consumer waste and material hotspots that are often hidden in supply chains.
        \item[Bullet 5: Future vision or the most important implications for continued research] Improved data availability and quality would enable more detailed and accurate waste and material footprinting.
    \end{description}

    \newpage
    % %% GRAPHICAL ABSTRACT
    % \begin{graphicalabstract}
    %     %\includegraphics{grabs}
    % \end{graphicalabstract}

    %% RESEARCH HIGHLIGHTS (3-5 bullet points, max 85 characters each)
    \begin{highlights}
        \item WasteAndMaterialFootprint, a new tool for quantifying waste and material flows in LCA.
        \item Assesses supply risks by calculating demand for critical materials.
        \item Simplifies quantification of user-specified waste and material categories.
        \item Rapidly identifies waste and material demand hotspots.
        \item Presents a case study of the battery supply chain.
    \end{highlights}

@@

    \title{WasteAndMaterialFootprint: A python package to quantify supply chain flows of waste and material in LCA databases}
    \author[1]{Stewart Charles McDowall\corref{cor1}}
    \ead{s.c.mcdowall@cml.leidenuniv.nl}
    \author[1]{Elizabeth Lanphear}
    \author[1]{Stefano Cucurachi}
    \author[1]{Carlos Felipe Blanco}

@@

    \begin{abstract}
        \cbox{Abstract word count: 150, Limit: 150}

        The `circular economy' is a concept centered on `re-X' strategies---on the prevention of waste and the reuse of resources. Thus, identifying and quantifying waste and material flows are of fundamental importance. Life Cycle Assessment (LCA) is powerful for this end, given its capacity to pinpoint hotspots of environmental impact throughout the life cycle of products and services, those where the implementation of circular principles could be most effective.

        Introducing WasteAndMaterialFootprint, a Python tool extending the Brightway framework for flexible, user-defined quantification of supply-chain demands in current and future scenarios. This tool streamlines database manipulation for LCA practitioners and integrates methods to aggregate and analyse waste and material flows, facilitating rapid hotspot identification.

        A case study on battery supply chains demonstrates the tool's utility. WasteAndMaterialFootprint quantifies and compares inventory demands and potential environmental burdens, aiding sustainable decision-making. It contributes to a more sustainable, circular economy by providing detailed material usage and waste generation analysis.

@@

\section{Conclusions}
\label{sec:conclusions}
\cbox{Section word count: 250}
We have written the WasteAndMaterialFootprint (WMF) tool, an extension to the brightway2 LCA framework that enables the user to calculate the waste and material footprints of a product or service in an LCA database. It explodes the database, identifies upstream waste and material exchanges, edits them, and writes matching custom WMF methods. These exchanges become pseudo-biosphere flows and thus, the footprint can be calculated as with the existing LCIA methods. The WMF tool can be easily customised by the user to calculate the footprints of other supply-chain flows such as water, gas, and critical raw materials.

@@

% Back matter
\section*{Data availability} 
All data used in this analysis are publicly available online under the noted sources. The WasteAndMaterialFootprint tool is installable via the Python Package Index (PyPI) and is available at \url{https://pypi.org/project/WasteAndMaterialFootprint}.
The full source code for the WasteAndMaterialFootprint tool is available at \url{https://www.github.com/Stew-McD/WasteAndMaterialFootprint}.
A user guide and comprehensive documentation are available at \url{https://wasteandmaterialfootprint.readthedocs.io}.

@@

\section*{Supplementary material}
    The supplementary material contains the following:
    \begin{enumerate}
        \item List of waste and material categories
        \item Example output of the WasteAndMaterialFootprint tool
        \item List of identified waste and material exchanges
        \item Code (python script) used for case study
        \item Inventory and methods used in the case study
        \item Complete tabulated results of the case study
        \item Complete visualisations of the case study
    \end{enumerate}

>>>>

    % SPOTLIGHTS (5 bullet points, max 120 characters each)
    {\Large Spotlights}
    \vspace{1em}
    \begin{description}[style=nextline]
        \item[Bullet 1: Critical context and background information on the problem addressed] Tracking waste and material flows is essential for developing the circular economy and supply chain resilience
        \item[Bullet 2: A brief overview of the key finding of the study (or findings if necessary)] A tool was developed to quantify waste and material footprints of activities in life cycle assessment (LCA) databases
        \item[Bullet 3: The most radical, creative, disruptive or innovative aspect of the manuscript] Built to complement the Brightway LCA framework and ActivityBrowser, the tool is customisable and easy to use.
        \item[Bullet 4: The significance of the results to the environment, economics or society] The tool can be used to identify pre-consumer waste and material hotspots that are often hidden in supply chains.
        \item[Bullet 5: Future vision or the most important implications for continued research] Improved data availability and quality would enable more detailed and accurate waste and material footprinting.
    \end{description}

    \newpage
    %% GRAPHICAL ABSTRACT
    \begin{graphicalabstract}
        \includegraphics{grabs}
    \end{graphicalabstract}

    %% RESEARCH HIGHLIGHTS (3-5 bullet points, max 85 characters each)
    \begin{highlights}
        \item T-reX, a new tool for quantifying waste and material flows in LCA.
        \item Assesses supply risks by calculating demand for critical materials.
        \item Simplifies quantification of user-specified waste and material categories.
        \item Rapidly identifies waste and material demand hotspots.
        \item Presents a case study of the battery supply chain.
    \end{highlights}

@@

    \title{T-reX: A python package to quantify supply chain flows of waste and material in LCA databases}
    \author[1]{Stewart Charles McDowall\corref{cor1}}
    \ead{s.c.mcdowall@cml.leidenuniv.nl}
    \author[1]{Elizabeth Lanphear}
    \author[1]{Stefano Cucurachi}
    \author[1]{Carlos Felipe Blanco}

@@

    \begin{abstract}
        \cbox{Abstract word count: 164, Limit: 150}
        Minimising waste through the reuse of resources is the quintessential principle of the `circular economy', but relies on our ability to identify and quantify waste and material flows.
        Thus, identifying and quantifying waste and material flows are of fundamental importance. Life Cycle Assessment (LCA) is powerful for this end, given its capacity to pinpoint hotspots of environmental impact throughout the life cycle of products and services, those where the implementation of circular principles could be most effective.

        Introducing T-reX, a Python tool extending the Brightway framework for flexible quantification of user-defined supply-chain demands in current and future scenarios. This tool streamlines database manipulation for LCA practitioners and integrates methods to aggregate and analyse waste and material flows, facilitating rapid hotspot identification.

        A case study on battery supply chains demonstrates the tool's utility. T-reX quantifies and compares inventory demands and, thus, potential environmental burdens, aiding sustainable decision-making. It contributes to the development of the `circular economy' by providing detailed material usage and waste generation analysis.

@@

\section{Conclusions}
\label{sec:conclusions}
\cbox{Section word count: 250}
We have written the T-reX tool, an extension to the brightway2 LCA framework that enables the user to calculate the waste and material footprints of a product or service in an LCA database. It explodes the database, identifies upstream waste and material exchanges, edits them, and writes matching custom WMF methods. These exchanges become pseudo-biosphere flows and thus, the footprint can be calculated as with the existing LCIA methods. The WMF tool can be easily customised by the user to calculate the footprints of other supply-chain flows such as water, gas, and critical raw materials.

@@

% Back matter
\section*{Data availability} 
All data used in this analysis are publicly available online under the noted sources. The T-reX tool is installable via the Python Package Index (PyPI) and is available at \url{https://pypi.org/project/T-reX}.
The full source code for the T-reX tool is available at \url{https://www.github.com/Stew-McD/T-reX}.
A user guide and comprehensive documentation are available at \url{https://T-reX.readthedocs.io}.

@@

\section*{Supplementary material}
    The supplementary material contains the following:
    \begin{enumerate}
        \item List of waste and material categories
        \item Example output of the T-reX tool
        \item List of identified waste and material exchanges
        \item Code (python script) used for case study
        \item Inventory and methods used in the case study
        \item Complete tabulated results of the case study
        \item Complete visualisations of the case study
    \end{enumerate}

### 2024-03-05 latex_files/methodology.tex
update abstract and introduction

This section is divided into two parts. In \autoref{sec:method-WMF_tool}, we describe the WasteAndMaterialFootprint (WMF) tool, in \autoref{sec:method-case_study}, we describe the methodology used to calculate the waste and material footprints in the case study.

\subsection{The WasteAndMaterialFootprint (WMF) tool}
\label{sec:method-WMF_tool}

@@

\begin{table}[h]
    \caption{WasteAndMaterialFootprint (WMF) tool metadata}
    \label{tab:methods-wmfmetadata}
    \centering
    \begin{tabular}{ll}
    \toprule
    \textbf{Item} & \textbf{Details} \\
    \midrule
    Current version & 0.1.21 \\
    DOI & \url{zenodo.org/doi/10.5281/zenodo.10431180} \\
    Code repository & \url{github.com/Stew-McD/WasteAndMaterialFootprint} \\
    License & CC0--1.0 license \\
    Versioning system & git \\
    Language & Python \\
    Documentation & \url{wasteandmaterialfootprint.readthedocs.io} \\
    Main dependencies & brightway2, premise, wurst \\
    Support email & s.c.mcdowall@cml.leidenuniv.nl \\
    \bottomrule
    \end{tabular}
\end{table}

@@

\subsubsection{Configuration modules}
\begin{itemize}
    \item \texttt{custom\_config}: Provides functions for managing the configuration of the WasteAndMaterialFootprint package.
    \item \texttt{user\_settings}: The main configuration file, for defining the project and database settings (user editable).
    \item \texttt{queries\_waste}: Defines search parameters and categories for waste generation exchanges (user editable).
    \item \texttt{queries\_materials}: Defines search parameters and categories for material demand exchanges (user editable).
\end{itemize}

\autoref{fig:WMF_flowchart} shows the flowchart of the WasteAndMaterialFootprint tool. The subsequent subsections describe the computational framework and the modules in more detail.

\begin{figure}[h!]
    \centering
    \includegraphics[height=0.8\textheight]{figures/WMF_flowchart.pdf}
    \caption{Flowchart of the WasteAndMaterialFootprint (WMF) tool}\label{fig:WMF_flowchart}
\end{figure}

@@

\subsubsection{Creation of custom `pseudo-biosphere' databases}
Custom `pseudo-biosphere' databases are created by \texttt{make\_custom\_database} module. This module collates the waste and material categories that were present in the databases, producing an \texttt{.xlsx} file that is imported back into the Brightway2 project as a biosphere-database named `WasteAndMaterialFootprint'.

>>>>

This section is divided into two parts. In \autoref{sec:method-WMF_tool}, we describe the T-reX tool, in \autoref{sec:method-case_study}, we describe the methodology used to calculate the waste and material footprints in the case study.

\subsection{The T-reX tool}
\label{sec:method-WMF_tool}

@@

\begin{table}[h]
    \caption{T-reX tool metadata}
    \label{tab:methods-wmfmetadata}
    \centering
    \begin{tabular}{ll}
    \toprule
    \textbf{Item} & \textbf{Details} \\
    \midrule
    Current version & 0.1.21 \\
    DOI & \url{zenodo.org/doi/10.5281/zenodo.10431180} \\
    Code repository & \url{github.com/Stew-McD/T-reX} \\
    License & CC0--1.0 license \\
    Versioning system & git \\
    Language & Python \\
    Documentation & \url{T-reX.readthedocs.io} \\
    Main dependencies & brightway2, premise, wurst \\
    Support email & s.c.mcdowall@cml.leidenuniv.nl \\
    \bottomrule
    \end{tabular}
\end{table}

@@

\subsubsection{Configuration modules}
\begin{itemize}
    \item \texttt{custom\_config}: Provides functions for managing the configuration of the T-reX package.
    \item \texttt{user\_settings}: The main configuration file, for defining the project and database settings (user editable).
    \item \texttt{queries\_waste}: Defines search parameters and categories for waste generation exchanges (user editable).
    \item \texttt{queries\_materials}: Defines search parameters and categories for material demand exchanges (user editable).
\end{itemize}

\autoref{fig:WMF_flowchart} shows the flowchart of the T-reX tool. The subsequent subsections describe the computational framework and the modules in more detail.

\begin{figure}[h!]
    \centering
    \includegraphics[height=0.8\textheight]{figures/WMF_flowchart.pdf}
    \caption{Flowchart of the T-reX tool}\label{fig:WMF_flowchart}
\end{figure}

@@

\subsubsection{Creation of custom `pseudo-biosphere' databases}
Custom `pseudo-biosphere' databases are created by \texttt{make\_custom\_database} module. This module collates the waste and material categories that were present in the databases, producing an \texttt{.xlsx} file that is imported back into the Brightway2 project as a biosphere-database named `T-reX'.
